Supporting Information for ``Effects of pre-existing structures on the seismicity of the Charlevoix Seismic Zone"

### 2019-07-01 Rebuttal_Letter.tex

  We have added the computational efficiency of our model in the model geometry section (Lines 131-137) of the manuscript. We did not add the computational efficiency to Table 1 because it does not change significantly among the models except for the SD65R25 model, which took about twice longer to run for the same model time. We do not clearly understand why this model is particularly slow but further analyzing computational efficiency is beyond the scope of this paper.

\end{response}

\begin{response}{3. Line 131-133: I do not understand why you set your boundary conditions as the differential stress reaches 706 MPa. Do you set a criterion for shear stress somewhere else? Please make it more clear.}

  The differential stress of 706 MPa at 10 km depth is based on Byerlee's law under the assumption of the absence of pore-fluid pressure. The full derivation of this value is now provided in the supplementary information. Shear stress was not considered as a stopping criterion. In addition, we have included a discussion on the possibility and some consequences of pore-fluid pressure at depth (Lines 538-549).

@@

   We have added a quantitative assessment of the correlation between the change in differential stress and the observed seismicity by determining the percentage of hypocenters in the region with $\Delta\sigma_{D}^\prime \ge$ 0.5\% without accounting for uncertainty in hypocenter relocation (Line 209-213). This metric is now used for describing the model results and comparing different models throughout the revised manuscript. We also added a discussion on the choice of the friction coefficient in Lines 512-518.

\end{response}

\begin{response}{8. Line 312: You need to justify the mesh size of 4 km in Section 2.1.}

We decided to run this particular group of models to acquire necessary information without increasing computational cost too much. Low-friction models, especially those with $\mu$ equal to 0.1 and 0.2, are very expensive due to slow convergence rate in the non-linear solution scheme. Running all the models in this group at the same mesh resolution of 4 km ensures intra-group consistency; and comparison with 2 km-resolution models is also justified because the descritization error does not dramatically increase as the resolution increases to 4 km (Fig. S1). This discussion is now included in Lines 118 to 127 of the revised manuscript.

### 2019-07-01 Supporting_Information.tex
Removed all comments and created a manuscript without trackchanges

This supplement presents the resolution test in our study, the derivation of the stopping criterion, effect of density ratio, cohesion, and moduli ratio on stress models.

@@

This resolution test shows that the error generally increases with mesh resolution (Fig. \ref{S1}). %For example, the model has an error of about 0.02 and 0.16 in the $\sigma_{D}$ and $\phi_{\max}$ in element size of 2 km relative to 1 km, respectively (Fig. \ref{S2}). The error values increase to 0.06 and 0.26 when the element size is 4 km. 
The errors in displacement components and differential stress in models with element sizes of 2 km is less than 5 \% of the corresponding values in the model with 1 km resolution. The errors are less than 10 \% for models with element sizes below 4 km. In our study, we will use an element size of 2 km for our fault models. However, due to expensive computation time, we will use an element size of 4 km to evaluate model parameters, e.g., coefficient of friction.
%
\begin{figure}[h!]
\setfigurenum{S1}
\includegraphics[width=25pc]{Figures/res_test_CSZ.png}
\caption{Errors in displacement components and differential stress ($\sigma_{D}$) for models with element sizes ranging from 2 to 8 km relative to a model with 1 km element size.}
\label{S1}
\end{figure}

@@

\noindent\textbf{Stopping criterion of differential stress at 10 km depth}
%\vspace{10mm} %5mm vertical space

@@

From $\Delta$ABC in Fig.~\ref{S2}, the illustration of the Mohr circle and the Byerlees law, we get
\[
 \sin \alpha = \frac{\sigma_{1}-\sigma_{3}}{2} / \left[\frac{250}{3} + \frac{\sigma_{1}+\sigma_{3}}{2} \right].
\]
%
%\[
% \sin \alpha = \frac{3(\sigma_{1}-\sigma_{3})}{3(\sigma_{1}+\sigma_{3})+500},
%\]
%
After simpligying and solving for $\sigma_{1}$, we ge the following expression:
\[
 \sigma_{1} = \frac{\sigma_{3}+[\sigma_{3}+\frac{500}{3}] \sin(\alpha)}{1-\sin \alpha}.
\]

Based on Byerlee's law, $\tau$  = 50 MPa + 0.6$\,\sigma_n$, under the assumption of absence of pore-fluid pressure along the rift faults, $\tan \alpha = 0.6$, hence, $\sin \alpha = 0.5145$. $\sigma_{3}$ corresponds to the lithostatic stress amounting to 250.0 MPa at 10 km depth for a density of 2.55 kg/m$^{3}$.
%\add[EC]{for a density of XXXX kg/m$^{3}$}. 
With these values and the above expression, we get $\sigma_{1}$ = 956.5 MPa.
%\annote[EC]{956.5}{Pay attention to the number of significant figures! I don't think our calculations can afford 7 or 8-digit precision so I rounded off the original value leaving 4 digits.} MPa. 
The differential stress at 10 km depth with no pore-fluid pressure, defined as $\sigma_{D} = \sigma_{1} - \sigma_{3}$, is 706.5 MPa.

@@

%\note[EC]{Turn the first long phrase into a sentence. Right now it reads like a figure caption but it shouldn't in a normal paragraph.} The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) for models with different impact structure-crust density ratio (Fig. \ref{S3}).

@@

\begin{figure}[ht]
\setfigurenum{S3}
\includegraphics[width=25pc]{Figures/effect_of_density.png}
\caption{
%\note[EC]{Symbol descriptions are missing. Include those in Fig. S4 caption.} 
Effect of impact structure-crust density ratio on the percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$) and the change in $\phi_{\max}$ ($\Delta \phi_{\max}^{\prime}$) at 10 km depth. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. The outline of the impact structure is represented as black circle. Each figure is 70 $\times$ 70 km.}
\label{S3}
\end{figure}

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in the SD70R100 model (Fig. \ref{S4}). The model exhibit the effects of the rift faults since the impact structure and the crust have the same elastic moduli. 
%\note[EC]{Turn the first long phrase into a sentence. Right now it reads like a figure caption but it shouldn't in a normal paragraph.} The percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$), and the change in $\phi_{\max}$ ($\Delta\phi_{\max}$) relative to the reference model (SNFR25) at various depths from the SD70R100 model (Fig. \ref{S4}).  
$\Delta\sigma_{D}^{\prime}$ shows increased $\sigma_{D}$ relative to SNFR25 throughout the entire length of the rift faults at and below 15 km depth, hence the model predicts earthquakes in those regions.

@@

\vspace{10mm} %5mm vertical space
%\pagebreak

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in a series of models in which the rift faults dip at 70$^{\circ}$ (i.e., SD70R25) but their cohesion ($C$) is varied from 0 to 30 MPa while the friction coefficient ($\mu$) is fixed at 0.3 (Fig.~\ref{S6}). We assume that the fault parameters are the same for each of the three rift faults and uniform throughout the fault surfaces. %Six models with 0, 3, 5, 10, 20 and 30 MPa are constructed (Fig. \ref{S6}).

%\vspace{10mm} %5mm vertical space

@@

%\vspace{10mm} %5mm vertical space

Magnitudes of $\Delta\sigma_D^{\prime}$ % and \annote[EC]{$\phi_{\max}$}{Fig. S6B doesn't support this.} 
decrease with increasing cohesion. $\Delta\sigma_D^{\prime}$ maps spatially overlap with the distribution of seismicity %for all the fault cohesions, especially well 
when the cohesion is less than 5 MPa with about 70\% of the hypocenters fall in the region with $\Delta\sigma_D^{\prime} \ge$ 0.5\% (Fig. \ref{S6}A). %The percentages of the hypocenters in the region with $\Delta\sigma_D^{\prime} \ge$ 0.5\% decrease to 61\% and 47\% for cohesion values of 20 and 30 MPa, respectively.
%\annote[EC]{, especially well when the cohesion is less than 5 MPa}{What makes you say so? Can't you use the new metric for this argument?} (Fig. \ref{S6}A). 
However, the models could not explain a few earthquakes in the SW corner of the crater, where $\Delta\sigma_D^{\prime}$ is negative. The $\Delta\sigma_D^{\prime}$ is not affected by a change in cohesion between 0 and 5 MPa. However, $\phi_{\max}$ approaches the reference no-fault model as the cohesion increases (Fig. \ref{S6}B). As cohesion increases, $\Delta\phi_{\max}$ decreases in value and the rotation is concentrated in the fault region outside the crater (Fig. \ref{S6}C). $\Delta\phi_{\max}$ is up to 10.8$^\circ$ when the cohesion is less than 5 MPa. We present models with zero cohesion in the main text to explain the spatial distribution of observed seismicity in the CSZ.

%\note[EC]{Why do you make remarks on $\mu$ mentioning cohesion only little in the paragraph for the effects of cohesion?} Figure \ref{S6} suggests that $\mu$ is an essential fault parameter needed to explain the seismicity of the CSZ because the value of $\mu$ affects the slope of the Mohr-Coulomb failure envelope. We present models with zero cohesion in the main text because to explain the spatial distribution of observed seismicity in the CSZ.

%\pagebreak

>>>>

This supplement presents the resolution test in our study, the derivation of the stopping criterion, effect of density ratio, moduli ratio, and cohesion on stress models.

@@

This resolution test shows that the error generally increases with mesh resolution (Fig. \ref{S1}). The errors in displacement components and differential stress in models with element sizes of 2 km is less than 5 \% of the corresponding values in the model with 1 km resolution. The errors are less than 10 \% for models with element sizes below 4 km. In our study, we will use an element size of 2 km for our fault models. However, due to expensive computation time, we will use an element size of 4 km to evaluate model parameters, e.g., coefficient of friction.
%
\begin{figure}[h!]
\setfigurenum{S1}
\includegraphics[width=25pc]{Figures/res_test_CSZ.png}
\caption{Errors in displacement components and differential stress ($\sigma_{D}$) for models with element sizes ranging from 2 to 8 km relative to a model with 1 km element size.}
\label{S1}
\end{figure}

@@

\noindent\textbf{Stopping criterion of differential stress at 10 km depth}

@@

From $\Delta$ABC in Fig.~\ref{S2}, the illustration of the Mohr circle and the Byerlees law, we get
\[
 \sin \alpha = \frac{\sigma_{1}-\sigma_{3}}{2} / \left[\frac{250}{3} + \frac{\sigma_{1}+\sigma_{3}}{2} \right].
\]
%
After simplifying and solving for $\sigma_{1}$, we ge the following expression:
\[
 \sigma_{1} = \frac{\sigma_{3}+[\sigma_{3}+\frac{500}{3}] \sin(\alpha)}{1-\sin \alpha}.
\]

Based on Byerlee's law, $\tau$  = 50 MPa + 0.6$\,\sigma_n$, under the assumption of absence of pore-fluid pressure along the rift faults, $\tan \alpha = 0.6$, hence, $\sin \alpha = 0.5145$. $\sigma_{3}$ corresponds to the lithostatic stress amounting to 250.0 MPa at 10 km depth for a density of 2.55 kg/m$^{3}$. With these values and the above expression, we get $\sigma_{1}$ = 956.5 MPa. The differential stress at 10 km depth with no pore-fluid pressure, defined as $\sigma_{D} = \sigma_{1} - \sigma_{3}$, is 706.5 MPa.

@@

@@

\begin{figure}[ht]
\setfigurenum{S3}
\includegraphics[width=25pc]{Figures/effect_of_density.png}
\caption{Effect of impact structure-crust density ratio on the percentage change in differential stress ($\Delta\sigma_{D}^{\prime}$), SH$_{\max}$ orientation ($\phi_{\max}$) and the change in $\phi_{\max}$ ($\Delta \phi_{\max}^{\prime}$) at 10 km depth. Earthquakes within 2 km of each depth slice are represented by black dots in \ref{S3}A. The outline of the impact structure is represented as a black circle. Each figure is 70 $\times$ 70 km.}
\label{S3}
\end{figure}

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in the SD70R100 model (Fig. \ref{S4}). The model exhibit the effects of the rift faults since the impact structure and the crust have the same elastic moduli. $\Delta\sigma_{D}^{\prime}$ shows increased $\sigma_{D}$ relative to SNFR25 throughout the entire length of the rift faults at and below 15 km depth, hence the model predicts earthquakes in those regions.

@@

@@

We compute $\Delta \sigma_{D}^{\prime}$, $\phi_{\max}$ and $\Delta \phi_{\max}^{\prime}$ in a series of models in which the rift faults dip at 70$^{\circ}$ (i.e., SD70R25) but their cohesion ($C$) is varied from 0 to 30 MPa while the friction coefficient ($\mu$) is fixed at 0.3 (Fig.~\ref{S6}). We assume that the fault parameters are the same for each of the three rift faults and uniform throughout the fault surfaces.

@@

Magnitudes of $\Delta\sigma_D^{\prime}$ decrease with increasing cohesion. $\Delta\sigma_D^{\prime}$ maps spatially overlap with the distribution of seismicity when the cohesion is less than 5 MPa with about 70\% of the hypocenters fall in the region with $\Delta\sigma_D^{\prime} \ge$ 0.5\% (Fig. \ref{S6}A). However, the models could not explain a few earthquakes in the SW corner of the crater, where $\Delta\sigma_D^{\prime}$ is negative. The $\Delta\sigma_D^{\prime}$ is not affected by a change in cohesion between 0 and 5 MPa. However, $\phi_{\max}$ approaches the reference no-fault model as the cohesion increases (Fig. \ref{S6}B). As cohesion increases, $\Delta\phi_{\max}$ decreases in value and the rotation is concentrated in the fault region outside the crater (Fig. \ref{S6}C). $\Delta\phi_{\max}$ is up to 10.8$^\circ$ when the cohesion is less than 5 MPa. We present models with zero cohesion in the main text to explain the spatial distribution of observed seismicity in the CSZ.
